Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-04-15 draft_p2.tex
Fine grained updates for part IV.B.2 Results GKTL

\subsubsection{Application of the \ac{GKTL} algorithm to extreme fluctuations of the time-averaged drag}
\label{sec:gktl_drag}

@@

We performed three experiments, corresponding to three different values of the parameter $k$.
In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.
Each experiment was run using $512$ Intel processors, each processor being assigned the computation of $N_c / 512 = 32$ clones.
The walltime for one experiment was roughly two days.

% Commentaire de la figure IS_GKTL
Figure~\ref{fig:illustration_IS} illustrates importance sampling achieved by means of the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process. This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\\ref{sec:pdfs}).
Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.

@@

For a fixed number of trajectories $N$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite
size effects harm the efficiency and accuracy of the selection procedure.
Typically, at each selection step, the majority of the clones are resampled, with the expection of only one or two.
This effect is illustrated in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
This is a consequence of the fact that most of the trajectories in the biased ensemble overlap, because the resampling of
discarded trajectories is based on a very small subset of the ensemble of $N$ trajectories.
Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the
bias $k$ is increased.
An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_max$ as
a function of the ensemble size $N$. 
See~\cite{ref1,ref2,ref3} for further details.

>>>>

\subsubsection{Application of the \ac{gktl} algorithm to extreme fluctuations of the time-averaged drag}
\label{sec:gktl_drag}

@@

We performed three experiments, corresponding to three different values of the parameter $k$.
In general the value of $k$ must be set empirically, unless a \textit{large deviation regime} $T_a \to \infty$ is verified~\cite{ref_ld_regime}.
In the latter case, the parameter $k$ can be related to the typical amplitude of the sampled fluctuations through the Gartner-Ellis theorem~\cite{GEref}.

% Commentaire de la figure IS_GKTL
Figure~\ref{fig:illustration_IS} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process. This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\ref{sec:pdfs}).
Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.

@@

For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
For $k \gtrapprox k_{max}$ the majority of the clones are resampled, with the expection of only one or two.
Because the resampling is based on a very small subset of the ensemble of $N_c$ trajectories, most of the trajectories in the biased ensemble overlap.
This effect is visible in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
See~\cite{ref1,ref2,ref3} for further details.
